---
output:
  pdf_document: 
    latex_engine: xelatex
    extra_dependencies: 
      - float
      - contour
header-includes:
 - \usepackage{xcolor}
 - \usepackage{graphicx}   
 - \usepackage{float}      
 - \usepackage{wrapfig}    
 - \newfontfamily\headingfont{RussoOne-Regular.ttf}
 - \setmainfont{Arial}
 - \pagenumbering{gobble}
 - \usepackage{geometry}
fontsize: 10pt
geometry: left=1in, right=1in, top=1in, bottom=0in
params:
  first: "Tanner"
  last: "Houck"
  sDate: "2024-03-01"
  eDate: "2024-09-30"
---
```{r setup, echo=FALSE, warning=FALSE, output=FALSE, include=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(scales)
library(knitr)
library(tinytex)
library(kableExtra)
library(gridExtra)
library(caret)
library(ggpubr)
library(baseballr)

# player <- playerid_lookup(params$last, params$first)
# player <- player|>
#   filter(!is.na(mlbam_id))

# game <- scrape_statcast_savant(
#   start_date = params$date,
#   end_date = params$date,
#   playerid = player$mlbam_id,
#   player_type = "pitcher"
# )

game <- read.csv("C:/Users/alexi/Downloads/savant_data.csv")

#Rearranging the name and dates
full_name <- paste(params$first, params$last)
yearS <- strsplit(params$sDate, "-")[[1]][1]
monthS <- strsplit(params$sDate, "-")[[1]][2]
dayS <- strsplit(params$sDate, "-")[[1]][3]
dateRealS <- paste0(monthS, "/", dayS, "/", yearS)
yearE <- strsplit(params$eDate, "-")[[1]][1]
monthE <- strsplit(params$eDate, "-")[[1]][2]
dayE <- strsplit(params$eDate, "-")[[1]][3]
dateRealE <- paste0(monthE, "/", dayE, "/", yearE)
dateRange <- paste0(dateRealS," - ", dateRealE)

#pitcher teams
pitcherTeam <- ifelse(game$inning_topbot[1]=="Top", game$home_team, game$away_team)
batterTeam <- ifelse(game$inning_topbot[1]=="Bot", "home", "away")

#logo
teamLogo <- paste0("C:/Users/alexi/OneDrive/Documents/Research/Logos/",pitcherTeam,".png")
```
\vspace*{-2.65cm}
\begin{minipage}[c]{0.2\textwidth}
    \hspace*{-2cm}
    \raisebox{-0.5cm}{\includegraphics[width=3cm]{C:/Users/alexi/OneDrive/Documents/Research/Logos/mlbLogo.png}}
\end{minipage}
\begin{minipage}[c]{0.6\textwidth}
    \begin{center}
        {\headingfont\fontsize{40}{30}\selectfont `r full_name`}\\
        \vspace{0.2cm}
        {\fontsize{10}{8}\selectfont 2024 vs. LHH}
    \end{center}
\end{minipage}
\begin{minipage}[c]{0.2\textwidth}
    \hspace*{2cm}  
    \raisebox{-0.5cm}{\includegraphics[width=1.8cm]{`r teamLogo`}}
\end{minipage}
```{r game scores, message=FALSE, echo= FALSE, include= FALSE, out.width= '100%'}
#keeping only relevant pitches
game<-game |>
  filter(!pitch_name %in% c('Other','Undefined'))|>
  filter(!is.na(plate_x) & !is.na(plate_z))|>
  mutate(Count = paste(balls, strikes, sep = "-"))

#creating / modifying variables for table output
game$release_speed = as.numeric(game$release_speed)
game$launch_speed = as.numeric(game$launch_speed)
game$release_spin_rate = as.numeric(game$release_spin_rate)
game$pfx_x = as.numeric(game$pfx_x)
game$pfx_z = as.numeric(game$pfx_z)
game$release_pos_z = as.numeric(game$release_pos_z)
game$release_pos_x = as.numeric(game$release_pos_x)
game$release_extension = as.numeric(game$release_extension)
game$plate_z = as.numeric(game$plate_z)
game$plate_x = as.numeric(game$plate_x)*-1
game$IVB = (game$pfx_z)*12
game$HB = (game$pfx_x)*12

#setting color pallete 
palette<-rep(c("4-Seam Fastball" = "violet", "Sinker" = "deepskyblue4", "Slider" = "lightblue1", "Changeup" = "forestgreen", "Curveball" = "violetred3", "Cutter" = "wheat", "Splitter" = "pink1", "Knuckleball" = "wheat3", "Undefined" = "black", "Other" = "gray52", "Sweeper" = "purple3", "Forkball"="cyan2", "Screwball"="green3", "Slurve"="maroon1", "Split-Finger" = "pink1", "Knuckle Curve" = "grey85"))

```

```{r line table, fig.pos = "hold_position", echo=FALSE, warning=FALSE, message=FALSE, fig.align="center", out.width='100%'}
# #pitch log for the game specified
# pitcherLogs <- daily_pitcher_bref(params, params$date)
# 
# pitcherLogs <- pitcherLogs|>
#   filter(Name == full_name)
# 
# #helping with calculations
# game$Single = ifelse(game$events == "single",1,0)
# game$Double = ifelse(game$events == "double",1,0)
# game$Triple = ifelse(game$events == "triple",1,0)
# game$HomeRun = ifelse(game$events == "home_run",1,0)
# game$Out = ifelse(game$events == "grounded_into_double_play" | game$events == "fielders_choice" | game$events == "force_out" | game$events == "field_out",1,0)
# game$RBOE = ifelse(game$events == "field_error",1,0)
# game$BB = ifelse(game$events == "walk",1,0)
# game$HBP = ifelse(game$description == "hit_by_pitch",1,0)
# game$K = ifelse(game$events == "strikeout",1,0)
# game$SAC = ifelse(game$events == "sacrifice",1,0)
# game$StrikesThrown = ifelse(game$description %in% c("called_strike", "swinging_strike", "foul", "hit_into_play", "swinging_strike_blocked"),1,0)
# game$BallsThrown = game$description %in% c("ball")
# game$PA = game$Single + game$Double + game$Triple + game$HomeRun + game$Out + game$RBOE +
#   game$BB + game$HBP + game$K + game$SAC
# 
# game$AB = game$Single + game$Double + game$Triple + game$HomeRun + game$Out + game$K
# 
# game$BF=ifelse(game$events!="" | game$events != '',1,0)
# game$"PitchCount"=ifelse(game$pitcher!="",1,0)
# 
# 
# #Create prelim table of basic stats
# game1=game|>
#   filter(balls==0 & strikes == 0)
# 
# game1$"firstpitchstrike"=ifelse(game1$description=="called_strike" | game1$description=="swinging_strike" |game1$description=="hit_into_play" |game1$description=="foul",1,0)
# 
# FirstPitch = sum(game1$pitchofpa)
# PitchCount = sum(game$PitchCount)
# 
# #first table
# gamescores = game |>
#   summarise(
#     "BF/P" = paste(sum(BF),sum(PitchCount), sep = "/"),
#     "Strikes/Balls" = paste(sum(StrikesThrown), sum(BallsThrown), sep = "/"),
#     IP = pitcherLogs$IP,
#     "H" = (sum(Single)+sum(Double)+sum(Triple)+sum(HomeRun)),
#     "HR" = sum(HomeRun),
#     ER = pitcherLogs$ER,
#     Ks = sum(K),
#     "BB" = sum(BB),
#     ERA = pitcherLogs$ERA,
#     WHIP = pitcherLogs$WHIP,
#     AVG = round((sum(Single)+sum(Double)+sum(Triple)+sum(HomeRun))/sum(AB), digits = 3),
#     OPS = ((round((sum(Single)+sum(Double)+sum(Triple)+sum(HomeRun)+sum(RBOE)+sum(BB)+sum(HBP))/sum(PA), digits = 3))+(round((sum(Single)+sum(Double)*2+sum(Triple)*3+sum(HomeRun)*4)/sum(AB), 
#                 digits = 3)))
#   )
# 
# 
#   
#   gamescores$"Strike%"=paste(round((sum(game$StrikesThrown)/PitchCount)*100, digits = 0),"%",
#     sep="")

```

```{r pitch breakdown, fig.pos = "hold_position", echo = FALSE,warning=FALSE, fig.align="center", out.width='100%'}

#assigning a new dataframe to manipulate
dat <- game
#View(dat)

#creating necessary variables in dat
dat$"PitchCount"=ifelse(dat$pitcher!="",1,0)
dat$"Pitch Type" = dat$pitch_name
dat$"IZ" = ifelse(dat$plate_z<=3.5 & dat$plate_z>=1.5 & dat$plate_x <= 0.83 & dat$plate_x >= -0.83,1,0)
dat$"Swing" = ifelse(dat$description == 'swinging_strike' | dat$description == 'foul' | dat$"description" == 'hit_into_play',1,0)
dat$"Whiff" = ifelse(dat$Swing == 1 & dat$description == 'swinging_strike',1,0)

#filtering dataframe to be what is displayed on table
pitchbreakdown <- dat|>
  group_by(`Pitch Type`)|>
  summarize(
    Count = sum(PitchCount),
    "Usage%" = paste(round(sum(PitchCount, na.rm = TRUE)/sum(nrow(dat), na.rm = TRUE)*100, 
                       digits = 0), "%",sep = ""),
    Velo = paste(round(quantile(release_speed,.25, na.rm=TRUE),
          digits = 0),round(quantile(release_speed,.75,na.rm = TRUE),digits = 0), sep="-"),
    Top = round(max(release_speed, na.rm=TRUE),digits = 2),
    Spin = round(mean(release_spin_rate, na.rm = TRUE), digits = 0), 
    IVB = round(mean(IVB, na.rm= TRUE), digits = 2),
    HB = -1*round(mean(HB, na.rm= TRUE), digits = 2),
    "Z%" = paste(round(sum(IZ, na.rm = TRUE)/sum(PitchCount, na.rm = TRUE)*100, 
                       digits = 0), "%",sep = ""),
    "Whiff%" = ifelse(sum(Swing, na.rm = TRUE) > 0,
      paste(100 * round(sum(Whiff, na.rm = TRUE) / sum(Swing, na.rm = TRUE), digits = 2), 
            '%', sep = ''),'0%')
  )|> arrange(desc(Count))
```

```{r break chart, echo= FALSE, warning=FALSE, message=FALSE}
#creating the break chart
breakChart <- ggplot(game, aes(x=-1*(pfx_x*12),y=pfx_z*12, color= pitch_name,))+
  labs(x="Horizontal Break", y= "Induced Vertical Break", color= "Pitch Type")+
  xlim(-30,30)+ylim(-30,30)+
  geom_hline(yintercept=0)+
  geom_vline(xintercept=0)+
  geom_point()+
  coord_fixed(ratio = 1)+
  scale_color_manual(values=palette)+
  theme_bw()+
  theme(legend.position='none')
```

```{r strike zone, echo= FALSE, message= FALSE, warning= FALSE, out.height= '85%', out.width= '85%'}
# creating strike zone
strikezone <- ggplot(data = game, aes(x= as.numeric(plate_x), y= as.numeric(plate_z)))+
    scale_color_manual(values=palette)+
    geom_density_2d_filled(alpha = 0.5, show.legend = FALSE)+
    scale_size(range = c(0.01,3))+
    geom_rect(xmin= -0.83,xmax= 0.83,ymin= 1.5,ymax= 3.5, color = "black", 
              fill = "transparent",size=1.1)+
    geom_rect(xmin= -1.2,xmax= 1.2,ymin= 1.2,ymax= 3.8, color = "black", 
              linetype = "dashed", fill = "transparent")+
    geom_segment(aes(x= -0.708, y= 0.15, xend= 0.708, yend= 0.15), size = 0.5, 
                 color = "black")+
    geom_segment(aes(x= -0.708, y= 0.3, xend= -0.708, yend= 0.15), size = 0.5,
                 color = "black")+
    geom_segment(aes(x= -0.708, y= 0.3, xend= 0, yend= 0.5), size = 0.5, color = "black")+
    geom_segment(aes(x= 0, y= 0.5, xend= 0.708, yend= 0.3), size = 0.5, color = "black")+
    geom_segment(aes(x= 0.708, y= 0.3, xend= 0.708, yend= 0.15), size = 0.5, 
                 color = "black")+
    coord_fixed(ratio = 1)+
    scale_x_continuous(limits = c(-2,2))+
    scale_y_continuous(limits = c(0,5))+
    theme_bw()+  
    coord_fixed(ratio = 1)+
    labs(color= "Pitch Type")+
    theme(axis.title = element_blank(), axis.ticks = element_blank(), 
          axis.text = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), legend.position = 'none')+
    ggtitle("Strike Zone")
```

```{r wOBA zone, echo= FALSE, message= FALSE, warning= FALSE, out.height= '85%', out.width= '85%'}
woba <- game |>
  mutate(
    zone_x = cut(plate_x, breaks = c(-2, -0.83, 0.83, 2), labels = c("Left", "Center", "Right")),
    zone_z = cut(plate_z, breaks = c(0, 1.5, 3.5, 5), labels = c("Low", "Middle", "High")),
    outer_zone_x = cut(plate_x, breaks = c(-1.2, -0.83, 0.83, 1.2), labels = c("Left_Outer", "Center_Outer", "Right_Outer")),
    outer_zone_z = cut(plate_z, breaks = c(1.2, 1.5, 3.5, 3.8), labels = c("Low_Outer", "Middle_Outer", "High_Outer"))
  ) |>
  group_by(zone_x, zone_z) |>
  summarize(avg_WOBA = mean(woba_value, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    x = case_when(
      zone_x == "Left" ~ -0.55,
      zone_x == "Center" ~ 0,
      zone_x == "Right" ~ 0.55
    ),
    y = case_when(
      zone_z == "Low" ~ 1.8,
      zone_z == "Middle" ~ 2.5,
      zone_z == "High" ~ 3.2
    )
  )|>
  mutate(avg_WOBA = ifelse(is.na(avg_WOBA), 0.000, avg_WOBA))

wobaZone <- ggplot() +
  geom_tile(data = woba, aes(x = x, y = y, fill = avg_WOBA), color = "blue", alpha = 0.5) +
  scale_fill_gradient(low = "blue", high = "red", name = "wOBA", labels = scales::label_number(accuracy = 0.001))+
  scale_size(range = c(0.01, 3)) +
  geom_rect(xmin = -0.84, xmax = 0.84, ymin = 1.4, ymax = 3.6, color = "black", 
            fill = "transparent", size = 1.1) +
  geom_segment(aes(x = -0.708, y = 0.15, xend = 0.708, yend = 0.15), size = 0.5, 
               color = "black") +
  geom_segment(aes(x = -0.708, y = 0.3, xend = -0.708, yend = 0.15), size = 0.5,
               color = "black") +
  geom_segment(aes(x = -0.708, y = 0.3, xend = 0, yend = 0.5), size = 0.5, color = "black") +
  geom_segment(aes(x = 0, y = 0.5, xend = 0.708, yend = 0.3), size = 0.5, color = "black") +
  geom_segment(aes(x = 0.708, y = 0.3, xend = 0.708, yend = 0.15), size = 0.5, 
               color = "black") +
  coord_equal() +
  scale_x_continuous(limits = c(-2, 2)) +
  scale_y_continuous(limits = c(0, 5)) +
  theme_bw() +  
  coord_fixed(ratio = 1) +
  labs(color = "Pitch Type") +
  theme(axis.title = element_blank(), axis.ticks = element_blank(), 
        axis.text = element_blank(),panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  ggtitle("Damage")

```

```{r velo distribution chart, echo= FALSE, message= FALSE, warning= FALSE, out.width= '50%', out.height= '50%'}
medians <- game|>
  group_by(pitch_name)|>
  summarize(median_release_speed = median(release_speed, na.rm = TRUE))

pitchDist <- ggplot(game, aes(x = release_speed, fill = pitch_name)) +
  geom_density(alpha = 0.5, color = NA) +
  facet_grid(pitch_name ~ ., scales = "free_x", switch = "both") +
  scale_fill_manual(values = palette) +
  geom_vline(data = medians, aes(xintercept = median_release_speed), 
             linetype = "dashed", color = "black") +
  labs(title = "Pitch Velocity Distribution", x = "Velocity (mph)", y= element_blank()) +
  theme(strip.text.y = element_blank(),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill="white", color = "black", size=0.1),
        panel.grid.minor = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks.x = element_line(color = "black"),
        axis.ticks.y = element_blank())
```

```{r extension mound, echo = FALSE, message = FALSE, warning=FALSE, fig.width= 10, fig.height=4}
#plot with release_extension and a mound
ext <- ggplot(game, aes(x = release_extension, y = release_pos_z)) +
  geom_point(aes(color = pitch_name)) +
  scale_color_manual(values = palette) +
  xlim(-2.5, 20) +
  ylim(0, 8) +
  geom_segment(aes(x = 0, y = 0.833, xend = 18, yend = 0), size = 1, color = "brown") +
  geom_segment(aes(x = 0, y = 0.833, xend = -2, yend = 0), size = 1, color = "brown") +
  geom_segment(aes(x = -2, y = 0, xend = 18, yend = 0), size = 0.4, color = "brown") +
  geom_vline(xintercept = mean(game$release_extension, na.rm = TRUE), linetype = "dashed", color = "black", size = 0.5) +
  annotate("text", x = 12.5, y = 6.12, 
           label = paste0("Avg Extension: ", round(mean(game$release_extension, na.rm = TRUE), 2)), color = "black") +
  labs(y= element_blank(), x = "Extension (feet)", color = "Pitch Type")+
  theme_bw()+ 
  theme(legend.position = 'bottom', axis.ticks = element_blank(), 
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.text.y = element_blank())

```

```{r colors, echo= FALSE, message= FALSE, warning= FALSE}
backgroundColor = "black"
textColor = "white"
if(pitcherTeam == "AZ"){
  backgroundColor = "#a71930"
  textColor = "#e3d4ad"
}else if(pitcherTeam == "ATL"){
  backgroundColor = "#13274f"
  textColor = "#ce1141"
}else if(pitcherTeam == "BAL"){
  backgroundColor = "#df4601"
  textColor = "black"
}else if(pitcherTeam == "BOS"){
  backgroundColor = "#192c55"
  textColor = "#bd3039"
}else if(pitcherTeam == "CHC"){
  backgroundColor = "#0e3386"
  textColor = "#cc3433"
}else if(pitcherTeam == "CHW"){
  textColor = "#c4ced4"
}else if(pitcherTeam == "CIN"){
  backgroundColor = "#c6011f"
}else if(pitcherTeam == "COL"){
  backgroundColor = "#33006f"
  textColor = "#c4ced4"
}else if(pitcherTeam == "CLE"){
  backgroundColor = "#1a2e5a"
  textColor = "#e31937"
}else if(pitcherTeam == "DET"){
  backgroundColor = "#182d55"
  textColor = "#f26722"
}else if(pitcherTeam == "HOU"){
  backgroundColor = "#002d62"
  textColor = "#f4911e"
}else if(pitcherTeam == "KCR"){
  backgroundColor = "#7bb2dd"
}else if(pitcherTeam == "LAA"){
  backgroundColor = "#ba0021"
  textColor = "#c4ced4"
}else if(pitcherTeam == "LAD"){
  backgroundColor = "#005A9C"
}else if(pitcherTeam == "MIA"){
  backgroundColor = "#41748D"
  textColor = "#EF3340"
}else if(pitcherTeam == "MIL"){
  backgroundColor = "#12284b"
  textColor = "#ffc52f"
}else if(pitcherTeam == "MIN"){
  backgroundColor = "#002B5C"
  textColor = "#D31145"
}else if(pitcherTeam == "NYM"){
  backgroundColor = "#002D72"
  textColor = "#FF5910"
}else if(pitcherTeam == "NYY"){
  backgroundColor = "#0C2340"
}else if(pitcherTeam == "OAK"){
  backgroundColor = "#003831"
  textColor = "#EFB21E"
}else if(pitcherTeam == "PHI"){
  backgroundColor = "#002d72"
  textColor = "#E81828"
}else if(pitcherTeam == "PIT"){
  backgroundColor = "#FDB827"
  textColor = "#27251F"
}else if(pitcherTeam == "SD"){
  backgroundColor = "#2F241F"
  textColor = "#FFC425"
}else if(pitcherTeam == "SFG"){
  backgroundColor = "#27251F"
  textColor = "#FD5A1E"
}else if(pitcherTeam == "SEA"){
  backgroundColor = "#005C5C"
  textColor = "#c4ced4"
}else if(pitcherTeam == "STL"){
  backgroundColor = "#C41E3A"
}else if(pitcherTeam == "TB"){
  backgroundColor = "#8FBCE6"
}else if(pitcherTeam == "TEX"){
  backgroundColor = "#C0111F"
}else if(pitcherTeam == "TOR"){
  backgroundColor = "#134A8E"
}else if(pitcherTeam == "WSH"){
  backgroundColor = "#AB0003"
}
```

```{r page 1.1, echo= FALSE, message= FALSE, warning= FALSE, fig.height= 3}
#pitch breakdown table
pitchbreakdown |>
    kbl(format = "latex", digits = 3, booktabs = TRUE, linesep = "") |>
    kable_paper("striped", html_font = "Comic Sans") |>
    row_spec(0, background = backgroundColor, color = textColor, bold = TRUE) |>
    kable_styling(latex_options = c("scale_down", "hold_position"), 
                  full_width = FALSE,font_size = 8,position = "center")

#break chart, k zone, and usage charts
ggarrange(breakChart, strikezone, wobaZone, ncol= 3, nrow=1, common.legend=FALSE, widths = c(1.2, 0.7, 1.2), heights = c(1.2, 1, 1))
```

```{r page 1.2, echo= FALSE, message= FALSE, warning= FALSE, fig.height= 3}
#pitch distribution chart
pitchDist
```

```{r page 1.3, echo= FALSE, message= FALSE, warning= FALSE, fig.height= 2.2}
#extension chart
ext
```
